---
output: 
    github_document:
        toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Banksy

<!-- badges: start -->
<!-- badges: end -->

Banksy is an R package that incorporates spatial information to cluster cells 
in a feature space (e.g. gene expression). Spatial information is incorporated 
by averaging the features of the k nearest neighbours to generate new 
'neighbour' features for a given cell. This is concatenated to the cell's own 
features to generate a combined feature matrix which is used for constructing a 
nearest neighbour network. Leiden clustering is used to obtain 
spatially-informed clusters.

## Installation

```{r, eval=F}
remotes::install_github("jleechung/Banksy")
```

Load the package:
```{r}
library(Banksy)
```

## Quick Start

### Input data

Inputs consist of an expression matrix and cell locations. Sample data provided
with the package:
```{r}
expr <- readRDS(system.file('/extdata/expression.rds', package = 'Banksy'))
locs <- readRDS(system.file('/extdata/locations.rds', package = 'Banksy'))
```

The gene expression matrix for cells should either be a `data.frame` or sparse 
matrix of class `dgCMatrix`:
```{r}
head(expr[,1:5])
```

The locations of each cell as a `data.frame`:
```{r}
head(locs)
```

### Spatial clustering

First, create a *BanksyObject* with the expression matrix and cell locations.
```{r}
bank <- BanksyObject(own.expr = expr,
                     cell.locs = locs)
```

Compute the neighbour matrix, and scale and normalize matrix. The options below
are defaults and listed for clarity.
```{r}
bank <- ComputeBanksy(bank,
                      normalizeColumns = TRUE,
                      normalizeColumnsTo = 100,
                      zScaleRows = TRUE,
                      zScaleBeforeAveraging = FALSE,
                      zScaleOwnAfterAveraging = TRUE,
                      zScaleNbrAfterAveraging = TRUE)
```

Obtain clusters for selected parameters.

- `lambda`. A mixing parameter from 0 to 1 which determines how much spatial 
   information is incorporated.  
- `resolution`. Leiden clustering resolution.  
- `kneighbours`. Number of k neighbours to use for constructing sNN.
```{r}
bank <- ClusterBanksy(bank, lambda = 0.25,
                            resolution = 1.2,
                            kneighbours = 30)
```

To explore the parameter space, one can provide sequences for the parameters:
```{r, eval=F}
## Not run
bank <- ClusterBanksy(bank, 
                      lambda = seq(0, 0.75, by = 0.25),
                      resolution = seq(0.8, 1.2, by = 0.2),
                      kneighbours = seq(30, 40, by = 10))
```

### Visualization

UMAP visualization:
```{r}
plotUMAP(bank, params = 'res1.2_lam0.25_k30', pt.size = 0.02)
```

Spatial plot:
```{r}
plotSpatialDims(bank, params = 'res1.2_lam0.25_k30', pt.size = 0.5)
```

### Subsetting

The *BanksyObject* can be subset by cells, features or metadata columns.
```{r}
bankSubset1 <- SubsetBanksy(bank,
                              dimx = sdimx > 1000 & sdimx < 4000,
                              dimy = sdimy > 2000 & sdimy < 3000,
                              cells = c(cell_364, cell_367, cell_384,cell_389),
                              features = sample(rownames(own.expr(bank)), 10))
head(bankSubset1)
```

```{r}
bankSubset2 <- SubsetBanksy(bank,
                            metadata = res1.2_lam0.25_k30 %in% 1:4)
plotSpatialDims(bankSubset2, params = 'res1.2_lam0.25_k30', pt.size = 0.5)
```

## Session information
```{r}
sessionInfo()
```

