---
output: 
    github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/jleechung/Banksy/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jleechung/Banksy/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/jleechung/Banksy/branch/main/graph/badge.svg?token=OZZK4EDVH9)](https://codecov.io/gh/jleechung/Banksy)
[![Maintenance](https://img.shields.io/badge/Maintained%3F-yes-green.svg)](https://github.com/jleechung/Banksy/graphs/commit-activity)
<!-- badges: end -->

## Overview

BANKSY is a method for clustering spatial transcriptomic data by augmenting the
transcriptomic profile of each cell with an average of the transcriptomes of its
spatial neighbors. By incorporating neighborhood information for clustering, 
BANKSY is able to

- improve cell-type assignment in noisy data
- distinguish subtly different cell-types stratified by microenvironment
- identify spatial zones sharing the same microenvironment

BANKSY is applicable to a wide array of spatial technologies (e.g. 10x Visium, 
Slide-seq, MERFISH) and scales well to large datasets. For more details on 
use-cases and methods, see the [preprint](). 

## Installation

The *Banksy* package can be installed via `remotes`:

```{r, eval=F}
remotes::install_github("prabhakarlab/Banksy", dependencies = TRUE)
```

## Documentation

Detailed description of *Banksy* functionality and example analyses are available
at the [package webpage](https://prabhakarlab.github.io/Banksy/). 

*Banksy* comes installed with [documentation](https://prabhakarlab.github.io/Banksy/reference/index.html) of main functions and their 
usage, along with several vignettes which detail different use cases:

* [Working with Banksy objects](https://prabhakarlab.github.io/Banksy/reference/banksy-object.html):
  Introduction to the *BanksyObject* class which serves as a container for 
  *Banksy*.  
  
* [Mouse hippocampus MERFISH dataset](https://prabhakarlab.github.io/Banksy/hippocampus-analysis.html):
  Illustrates a grid search of parameters which best cluster cells.  

* [Human dorsolateral prefrontal cortex 10x Visium dataset](https://prabhakarlab.github.io/Banksy/dlpfc-analysis.html):
  Illustrates analysis of multiple spatial transcriptomic datasets.

* [Mouse hypothalamus MERFISH dataset](https://prabhakarlab.github.io/Banksy/hypothalamus-analysis.html):
  Illustrates visualization functionality.
  
* [Interoperability with SingleCellExperiment](https://prabhakarlab.github.io/Banksy/single-cell-exp.html):
  Illustrates BANKSY interoperability with the Bioconductor
  [SingleCellExperiment](https://bioconductor.org/packages/release/bioc/html/SingleCellExperiment.html)
  framework for interfacing with packages like [scran](https://bioconductor.org/packages/release/bioc/html/scran.html) 
  or [scater](https://bioconductor.org/packages/release/bioc/html/scater.html).

## Quick start

*Banksy* takes as input an expression matrix and cell centroids. Example 
datasets are provided with the package:

```{r, eval = T}
library(Banksy)

data(hippocampus)
expr <- hippocampus$expression
locs <- hippocampus$locations
```

The gene expression matrix for cells should be a `matrix`:

```{r, eval = T}
class(expr)
head(expr[,1:5])
```

while cell locations should be supplied as a `data.frame`:

```{r, eval = T}
class(locs)
head(locs)
```

We store the total counts for each cell and the number of expressed genes as 
metadata `data.frame`, which can optionally be supplied:

```{r, eval = T}
total_count <- colSums(expr)
num_genes <- colSums(expr > 0)
meta <- data.frame(total_count = total_count, num_genes = num_genes)
```

Next, create a *BanksyObject* with the expression matrix and cell locations 
(metadata is optional).

```{r, eval = T}
bank <- BanksyObject(own.expr = expr,
                     cell.locs = locs,
                     meta.data = meta)
```

Apply basic QC by keeping only cells with total counts within the 5th and 98th 
percentile:

```{r, eval = T}
bank <- SubsetBanksy(bank, metadata = total_count > quantile(total_count, 0.05) &
                                      total_count < quantile(total_count, 0.98))
```

We first normalize the expression matrix, compute the neighbor matrix, and 
scale the resulting matrices.

```{r, eval = T}
bank <- NormalizeBanksy(bank, normFactor = 100)
bank <- ComputeBanksy(bank, k_geom = 10, spatialMode = 'kNN_r')
bank <- ScaleBanksy(bank)
```

Run PCA on the BANKSY matrix for `lambda = 0` (no spatial information) 
and `lambda = 0.3`. 

```{r, eval = T}
bank <- RunPCA(bank, lambda = c(0, 0.3), npcs = 30)
```

Next, we obtain cluster assignments using graph-based clustering with the Leiden
algorithm on the first 20 PCs. Specify the following parameters:

* `resolution`. Leiden clustering resolution.  
* `k.neighbours`. Number of k neighbours to use for constructing sNN.

```{r, eval = T}
set.seed(42)
bank <- ClusterBanksy(bank, lambda = c(0, 0.3), pca = TRUE, npcs = 20,
                      method = 'leiden', k.neighbors = 50, resolution = 1.2)
```

Different clustering runs can be harmonised with `ConnectClusters`:

```{r, eval = T}
bank <- ConnectClusters(bank, map.to = clust.names(bank)[1])
``` 

Visualise the clustering output for non-spatial clustering (`lambda=0`) and 
BANKSY clustering (`lambda = 0.3`). 

```{r, eval = T, fig.height=8, fig.width=18}
features <- clust.names(bank)
feature.types <- rep('discrete', 2)
main <- c('Non-spatial', 'BANKSY')
plotSpatialFeatures(bank, by = features, type = feature.types, main = main, 
                    pt.size = 1.5, main.size = 15, nrow = 1, ncol = 2)
```

For clarity, we can visualise each of the clusters separately with `wrap = TRUE`:

```{r, eval = T, fig.height=8, fig.width=18}
plotSpatialFeatures(bank, by = features, type = feature.types, main = main, 
                    pt.size = 0.5, main.size = 15, nrow = 1, ncol = 2, 
                    wrap = TRUE)
``` 

## Contributing

Bug reports, questions, request for enhancements or other contributions can be 
raised at the [issue page](https://github.com/prabhakarlab/Banksy/issues). 
