---
title: "Clustering via BANKSY matrix construction"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Clustering via BANKSY matrix construction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = 'figures/'
)
```

By default, `runBanksyPCA` uses a lazy linear operator that computes PCA
without materializing the full BANKSY matrix. This is memory-efficient and
recommended for most use cases.

However, some workflows require explicit construction of the BANKSY matrix —
for example, when using the azimuthal Gabor filter (AGF), when access to the
intermediate matrices is needed, or when using higher harmonics (`M > 0`).
This vignette demonstrates the standard workflow using `computeBanksy`
followed by `runBanksyPCA(lazy=FALSE)`.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(Banksy)

library(SummarizedExperiment)
library(SpatialExperiment)
library(scuttle)

library(scater)
library(cowplot)
library(ggplot2)
```

# Data preprocessing

We use the mouse hippocampus dataset provided with *Banksy*.

```{r, eval=TRUE}
data(hippocampus)
gcm <- hippocampus$expression
locs <- as.matrix(hippocampus$locations)
```

Initialize a SpatialExperiment object and perform quality control and
normalization.

```{r, eval=TRUE, message=FALSE}
se <- SpatialExperiment(assay = list(counts = gcm), spatialCoords = locs)

# QC based on total counts
qcstats <- perCellQCMetrics(se)
thres <- quantile(qcstats$total, c(0.05, 0.98))
keep <- (qcstats$total > thres[1]) & (qcstats$total < thres[2])
se <- se[, keep]

# Normalization to mean library size
se <- computeLibraryFactors(se)
aname <- "normcounts"
assay(se, aname) <- normalizeCounts(se, log = FALSE)
```

# Computing the BANKSY neighborhood matrices

`computeBanksy` computes the neighborhood matrices and stores them as assays
in the SpatialExperiment object. Setting `compute_agf=TRUE` computes both
the weighted neighborhood mean ($\mathcal{M}$) and the azimuthal Gabor filter
($\mathcal{G}$). The number of spatial neighbors used to compute
$\mathcal{M}$ and $\mathcal{G}$ are `k_geom[1]=15` and `k_geom[2]=30`
respectively.

```{r, eval=TRUE}
k_geom <- c(15, 30)
se <- computeBanksy(se, assay_name = aname, compute_agf = TRUE, k_geom = k_geom)
```

After running `computeBanksy`, the neighborhood matrices are accessible as
assays:

```{r, eval=TRUE}
assayNames(se)
```

# PCA, clustering, and visualization

We run PCA with `lazy=FALSE` to use the standard path, which constructs the
full BANKSY matrix via `getBanksyMatrix`. Setting `use_agf=TRUE` uses both
$\mathcal{M}$ and $\mathcal{G}$. We compare non-spatial clustering
(`lambda=0`) with BANKSY cell-typing (`lambda=0.2`).

```{r, eval=TRUE}
lambda <- c(0, 0.2)
set.seed(1000)
se <- runBanksyPCA(se, use_agf = TRUE, lambda = lambda, lazy = FALSE)
se <- runBanksyUMAP(se, use_agf = TRUE, lambda = lambda)
se <- clusterBanksy(se, use_agf = TRUE, lambda = lambda, resolution = 1.2)
se <- connectClusters(se)
```

Visualise the clustering output:

```{r, eval=TRUE, fig.height=5, fig.width=14}
cnames <- colnames(colData(se))
cnames <- cnames[grep("^clust", cnames)]
colData(se) <- cbind(colData(se), spatialCoords(se))

plot_nsp <- plotColData(se,
    x = "sdimx", y = "sdimy",
    point_size = 0.6, colour_by = cnames[1]
)
plot_bank <- plotColData(se,
    x = "sdimx", y = "sdimy",
    point_size = 0.6, colour_by = cnames[2]
)

plot_grid(plot_nsp + coord_equal(), plot_bank + coord_equal(), ncol = 2)
```

# Accessing the BANKSY matrix

The full BANKSY matrix can be retrieved with `getBanksyMatrix`. This is
useful for custom downstream analyses.

```{r, eval=TRUE}
banksy_matrix <- getBanksyMatrix(se, M = 1, lambda = 0.2,
                                  assay_name = aname, scale = TRUE)
dim(banksy_matrix)
```

<details>
    <summary>Session information</summary>

```{r, sess}
sessionInfo()
```

</details>
