---
title: "Domain segmentation (STARmap PLUS mouse brain)"
output: rmarkdown::html_vignette
# output: pdf_document
vignette: >
  %\VignetteIndexEntry{Domain segmentation (STARmap PLUS mouse brain)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here, we demonstrate BANKSY domain segmentation on a STARmap PLUS dataset of 
the mouse brain from [Shi et al. (2022)](https://doi.org/10.1101/2022.06.20.496914). 

```{r, eval=T, include=F}
start.time = Sys.time()
```

```{r, eval=T, message=FALSE, warning=FALSE}
library(Banksy)

library(data.table)
library(SummarizedExperiment)
library(SpatialExperiment)

library(scater)
library(cowplot)
library(ggplot2)
```

## Data preprocessing

Data from the study is available from the [Single Cell Portal](https://singlecell.broadinstitute.org/single_cell/study/SCP1830). We 
analyze data from `well11`. The data comprise 1,022 genes profiled at 
subcellular resolution in 43,341 cells. 

```{r, eval=F}
#' Change paths accordingly
gcm_path = '../data/well11processed_expression_pd.csv.gz'
mdata_path = '../data/well11_spatial.csv.gz'

#' Gene cell matrix
gcm = fread(gcm_path)
genes = gcm$GENE
gcm = as.matrix(gcm[,-1])
rownames(gcm) = genes

#' Spatial coordinates and metadata
mdata = fread(mdata_path, skip=1)
headers = names(fread(mdata_path, nrows=0))
colnames(mdata) = headers
#' Orient spatial coordinates
xx = mdata$X
yy = mdata$Y
mdata$X = max(yy) - yy
mdata$Y = max(xx) - xx
mdata = data.frame(mdata)
rownames(mdata) = colnames(gcm)

locs = as.matrix(mdata[,c('X','Y','Z')])

#' Create SpatialExperiment
se <- SpatialExperiment(
    assay = list(processedExp = gcm),
    spatialCoords = locs,
    colData = mdata
)
```

## Running BANKSY

Run BANKSY in domain segmentation mode with `lambda=0.8`. This places larger
weights on the mean neighborhood expression and azimuthal Gabor filter in 
constructing the BANKSY matrix. 

```{r, eval=F}
lambda <- 0.8
k_geom <- 30
npcs <- 50
aname <- 'processedExp'
se <- Banksy::computeBanksy(se, assay_name = aname, k_geom = k_geom)

set.seed(1000)
se <- Banksy::runBanksyPCA(se, lambda = lambda, npcs = npcs)
```

We adjust the resolution to yield 23 clusters based on the results from 
[Maher et al. (2023)](https://doi.org/10.1101/2023.06.30.547258v1) 
(see Fig. 1, 2).  

```{r, eval=F}
set.seed(1000)
se <- Banksy::clusterBanksy(se, lambda = lambda, npcs = npcs, resolution = 0.8)
```

Visualize clustering results:

```{r, eval=T, echo=F}
se <- readRDS(system.file('extdata/STARmap.rds', package='Banksy'))
```

```{r, eval=T, fig.height=8, fig.width=7, fig.align='center'}
cnames <- colnames(colData(se))
cnames <- cnames[grep('^clust', cnames)]

plotColData(se, x = 'X', y = 'Y', point_size = 0.01, colour_by = cnames[1]) +
    scale_color_manual(values = pals::glasbey()) + 
    coord_equal() +
    theme(legend.position = 'none')
```

## Session information

<details>

```{r, sess}
sessionInfo()
```

</details>
