---
title: "Mouse Hippocampus VeraFISH analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mouse Hippocampus VeraFISH analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here, we demonstrate a grid search of clustering parameters with a mouse 
hippocampus VeraFISH dataset. *BANKSY* currently provides four algorithms for 
clustering the BANKSY matrix with *ClusterBanksy*: Leiden (default), Louvain,
k-means, and model-based clustering. In this vignette, we run only Leiden 
clustering. See `?ClusterBanksy` for more details on the parameters for
different clustering methods.

## Loading the data

First, we construct the *BanksyObject*. The dataset comprises gene expression 
for 10,944 cells and 120 genes in 2 spatial dimensions. See 
`?Banksy::hippocampus` for more details.

```{r load, eval=T, message=F, warning=F}
library(Banksy)
library(ggplot2)
library(gridExtra)

data(hippocampus)
expr <- hippocampus$expression
locs <- hippocampus$locations

# Store total counts and number of expressed genes
total_count <- colSums(expr)
num_genes <- colSums(expr > 0)
meta <- data.frame(total_count = total_count, num_genes = num_genes)

# Construct BanksyObject
bank <- BanksyObject(own.expr = expr, cell.locs = locs, meta.data = meta)
bank
```

## Parameters

The `lambda` parameter is a mixing parameter in `[0,1]` which 
determines how much spatial information is incorporated for clustering. With 
smaller values of `lambda` (e.g. `[0,0.25]`), BANKY operates in 
*cell-typing* mode, while at higher levels of `lambda`, BANKSY operates in 
*zone-finding* mode. See the pre-print for more details.

Leiden graph-based clustering admits two parameters: `k.neighbors` and 
`resolution`. `k.neighbors` determines the number of k nearest neighbors used to
construct the shared nearest neighbors graph. Leiden clustering is then 
performed on the resultant graph with resolution `resolution`.

We first process the data for clustering:

```{r process, eval=T, message=F, warning=F}
bank <- SubsetBanksy(bank, metadata = total_count > quantile(total_count, 0.05) &
                                      total_count < quantile(total_count, 0.98))
bank <- NormalizeBanksy(bank)
bank <- ComputeBanksy(bank)
bank <- ScaleBanksy(bank)
```

## Grid search

Before clustering, we define a parameter space. Here, we'll explore different 
combinations of `lambda` and `resolution` while fixing `kneighbours`. We first
run PCA on the BANKSY matrix with 20 PCs for each `lambda`. Clustering is then
performed on the PCs (by default):

```{r params, eval=T}
lam <- c(0, 0.3)
res <- c(0.8, 1.2)
knbr <- 50

bank <- RunPCA(bank, lambda = c(0, 0.3), npcs = 20)
set.seed(42)
bank <- ClusterBanksy(bank, lambda = lam, pca = TRUE, npcs = 20, 
                      method = 'leiden', k.neighbors = knbr, resolution = res)
```

This populates the `meta.data` slot of the *BanksyObject* with cluster labels
for each combination of parameters:

```{r mdata, eval=T}
head(meta.data(bank))
```

## Cluster similarity

To compare clustering labels from different parameter combinations, the package
implements the *getARI* and *plotARI* functions. These compute the adjusted 
Rand index (ARI) for all pairs of clusters to assess the similarity of 
clustering labels. Observe that clusters with the same `lambda` share more
similarity:

```{r runari, eval=T, fig.height=4.5, fig.width=5, fig.align='center'}
getARI(bank)

plotARI(bank) 
```

## Connecting clusters

To visually compare between clusters obtained with different combination of 
parameters (a parameter run), the package implements *ConnectClusters* which
performs a mapping to harmonise cluster labels between different parameter runs.

```{r connect, eval=T}
bank <- ConnectClusters(bank, map.to = 'clust_lam0.3_k50_res1.2')

head(meta.data(bank))
```

This updates the `meta.data` slot with new cluster labels that can be used for 
visualisation. We can visualise connected output as follows.

First, obtain cluster names:

```{r clustnames, eval=T}
cnms <- clust.names(bank)
```

Then, visualise the cluster runs:

```{r out, eval=T, fig.width=10, fig.height=8}
plotSpatialFeatures(bank, by = cnms, type = rep('discrete', 4), 
                    nrow = 2, ncol = 2, main = cnms, main.size = 10)
```

We visualise the differences between clusters obtained with non-spatial 
clustering (`lam=0`) and BANKSY in cell-typing mode (`lam=0.3`) more finely by 
splitting the clusters with `wrap = TRUE`:

```{r wrap, eval=T, fig.width=12, fig.height=6, fig.align='center'}
p1 <- plotSpatial(bank, by = cnms[2], type = 'discrete', 
            main = cnms[2], main.size = 10, pt.size = 0.1, wrap = TRUE)

p2 <- plotSpatial(bank, by = cnms[4], type = 'discrete', 
            main = cnms[4], main.size = 10, pt.size = 0.1, wrap = TRUE)

grid.arrange(p1, p2, ncol = 2)
```

In addition, one can also visualise differences in the clustering labels with
alluvial plots:

```{r alluv, eval=T, fig.width=7, fig.height=6, fig.align='center'}
plotAlluvia(bank)
```

## Session information

<details>

```{r, sess}
sessionInfo()
```

</details>

