---
title: "Visualisation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualisation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This article demonstrates functions for visualisation with a mouse hypothalamus
MERFISH dataset. The dataset comprises 11,162 cells and 161 genes in 3 spatial 
dimensions. Cell type annotation along with other metadata is provided. 

```{r, init, eval=TRUE}
library(Banksy)

data(hypothalamus)

# Initialize BanksyObject
expr <- hypothalamus$expression
locs <- hypothalamus$locations
meta <- hypothalamus$metadata

total_count <- colSums(expr)
meta <- cbind(meta, total_count = total_count)

bank <- BanksyObject(own.expr = expr,
                     cell.locs = locs,
                     meta.data = meta)

# Filter BanksyObject based on total count
bank <- SubsetBanksy(bank, metadata = total_count > quantile(total_count, 0.05))

bank
``` 

Almost all plotting functions in *Banksy* return *ggplot* objects. They can thus
be further modified with the grammar of *ggplot*.

## Spatial plots

*Banksy* implements `plotSpatial` for visualising cells in spatial dimensions. 
For 3 dimensional datasets, it is assumed that the z-plane is
discretized. Spatial plots will be wrapped by z-plane. 

```{r, spatial-1, fig.width=7, fig.height=3.5, fig.align='center'}
plotSpatial(bank)
``` 

Cells can be colored by groups. Any column in the metadata and any feature 
(gene) can be used as a grouping variable.

```{r, spatial-2, fig.width=7, fig.height=3.5, fig.align='center'}
plotSpatial(bank, by = 'Cell_class', type = 'discrete')
```

Plots can be wrapped with `wrap=TRUE`:

```{r, wrap, fig.width=14, fig.height=4, fig.align='center'}
# Wrap to visualise each group in the grouping factor
plotSpatial(bank, by = 'Cell_class', type = 'discrete', wrap = TRUE)
``` 

Process the data:

```{r, process}
bank <- NormalizeBanksy(bank)
bank <- ComputeBanksy(bank, k_geom = 15, spatialMode = 'kNN_r')
bank <- ScaleBanksy(bank)
``` 

Visualise scaled *Mbp* expression, a marker for myelinating oligodendrocytes:

```{r, spatial-3, fig.width=7, fig.height=3.5, fig.align='center'}
plotSpatial(bank, by = 'Mbp', type = 'continuous', 
            col.lowpoint = -1, col.highpoint = 1)
``` 

Multiple features can be plotted simultaneously with `plotSpatialFeatures`:

```{r, spatial-4, fig.width=7, fig.height=10.5, fig.align='center'}
features <- c('Mbp', 'Cxcl14', 'Fn1')
type <- rep('continuous', 3)
plotSpatialFeatures(bank, by = features, type = type, 
                    col.lowpoint = -1, col.highpoint = 1, 
                    main = features, main.size = 10, nrow = 3)
``` 

## Dimensionality reduction 

*Banksy* provides functions for running and visualising principal component 
analysis (PCA) and uniform manifold approximation projection (UMAP) for 
dimensionality reduction.

We run PCA and UMAP, and visualise the results. A scree plot can be used to 
visualise the proportion of variance explained by each PC, and determine
how many PCs are used in downstream analyses.

```{r, pca, fig.width=7, fig.height=3.5, fig.align='center'}
bank <- RunPCA(bank, lambda = 0.2, npcs = 30)
plotReduction(bank, reduction = 'pca_0.2', by = 'Cell_class', type = 'discrete')
plotScree(bank, lambda = 0.2)
```

Next, we run UMAP and visualise the projection:

```{r, umap, fig.width=7, fig.height=3.5, fig.align='center'}
bank <- RunUMAP(bank, lambda = 0.2, npcs = 20)
p1 <- plotReduction(bank, reduction = 'umap_0.2', by = 'Cell_class', 
                    type = 'discrete', pt.size = 0.25)
```

Run BANKSY with Leiden clustering, and compare the clustering output with the
cell class annotation:

```{r, cluster, fig.width=9, fig.height=4, fig.align='center'}
bank <- ClusterBanksy(bank, lambda = 0.2, pca = TRUE, npcs = 20,
                      method = 'leiden', k.neighbors = 50, resolution = 1)
p2 <- plotReduction(bank, reduction = 'umap_0.2', by = clust.names(bank), 
                    type = 'discrete', pt.size = 0.25)

gridExtra::grid.arrange(
    p1, p2, ncol = 2
)
```

## Heatmaps

We implement heatmap visualisation with the *ComplexHeatmap* package. The 
`assay` argument takes one of:

- `own.expr`: visualise the cell's own expression
- `nbr.expr`: visualises the neighborhood expression
- `banksy`: visualises combined own and neighbor expression weighted by `lambda`

We can introduce multiple cell annotations by setting `annotate=TRUE` and 
specifying `annotate.by`. 

Because expression matrices can be large, one can also specify the maximum 
number of columns that should be plotted with `max.cols`. This subsamples the
matrix to speed up plotting.

```{r, heatmap, fig.width=9, fig.height=6, fig.align='center'}
set.seed(1000)
plotHeatmap(bank, assay = 'banksy', lambda = 0.2, 
            annotate = TRUE, 
            annotate.by = c('Cell_class', clust.names(bank)), 
            order.by = clust.names(bank), 
            features = sample(rownames(own.expr(bank)), 100),
            max.cols = 200)
```

## Session information

```{r}
sessionInfo()
```

