---
title: "Varying Neighboughood Size and Kernel Shape"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Varying Neighboughood Size and Kernel Shape}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning=  FALSE,
  collapse = TRUE,
  comment = "#>"
)
#pkgdown::build_site_github_pages()
```

In this vignette, we explore how variation in the size of the spatial neighbourhood affects the clustering results. Recall that a cell's microenvironment signature is computed as a weighted average of the expression of that cell's neighbours. This weighting can be done using a kernel, such as $1/r^n$ for $n \in \{1, 2\}$, where $r$ is the distance of a given neighbour cell from the index cell. Other kernels include the Gaussian kernel, or the rank-Gaussian kernel; see the manuscript for details. 

## Load packages, data and set parameters
```{r, echo=TRUE, eval=TRUE}
library(Banksy)
library(ggplot2)
library(gridExtra)
library(scales)
library(ComplexHeatmap)
library(data.table)
library(irlba)
library(MERINGUE)
library(Matrix)

# load data
data(hippocampus)
expr <- hippocampus$expression
locs <- hippocampus$locations

k_geom_list = rep(c(5, 10, 20, 50, 100)) # number of cells in spatial neighbourhood
num_nbd = 5 # set to an integer between 1 and 5 to select which elements of k_geom_list to try. 
# For eg., with num_nbd = 2, we try k_geom = 5 and 10. 

```


## Process data and perform clustering with different neighbourhood kernels
```{r}
subsamp = FALSE # set to FALSE if you don't want to subsample cells. 
total_count <- colSums(expr)
num_genes <- colSums(expr > 0)
meta <- data.frame(total_count = total_count, num_genes = num_genes)
bank_raw <- BanksyObject(own.expr = expr, cell.locs = locs, meta.data = meta)
if (subsamp){
   cells_sample = sample(bank_raw@meta.data$cell_ID, 1000)
} else {
  cells_sample = bank_raw@meta.data$cell_ID
}
bank_raw <- SubsetBanksy(bank_raw, cells = cells_sample,
                     metadata = total_count > quantile(total_count, 0.05) &
                       total_count < quantile(total_count, 0.98))
bank_raw <- NormalizeBanksy(bank_raw)

clusterDiffNbdSizes<-function(bank, spatialMode_list, k_geom_list){
  bankc<-copy(bank)
  clustering_results<-function(spatialMode, k_geom){
    bankc <- ComputeBanksy(bank=bankc, spatialMode = spatialMode, k_geom = k_geom)
    bankc <- ScaleBanksy(bankc)
    bankc <- Banksy:::RunPCA(bankc, lambda = 0.25, npcs = 20)
    set.seed(42)
    bankc <- ClusterBanksy(bankc, lambda = 0.25, pca = TRUE, npcs = 20,
                          method = 'leiden', k.neighbors = 50, resolution = 1.5)
    return(bankc@meta.data$clust_lam0.25_k50_res1.5) 
  }
  clustering_list <- mapply(clustering_results, spatialMode_list, k_geom_list)
  clustering_df<-as.data.frame(clustering_list)
  colnames(clustering_df)<-paste0('clust_lam0.25_k50_res1.5', '_', spatialMode_list, '_kg', k_geom_list)
  return(clustering_df)
}

bank <- copy(bank_raw) 

clust_kNN_r <- clusterDiffNbdSizes(bank = bank, spatialMode_list = rep('kNN_r', num_nbd), 
                                  k_geom_list = k_geom_list[1:num_nbd])
clust_kNN_r2 <- clusterDiffNbdSizes(bank = bank, spatialMode_list = rep('kNN_rn', num_nbd), 
                                  k_geom_list = k_geom_list[1:num_nbd])

# default banksy object
bank <- ComputeBanksy(bank, spatialMode = 'kNN_r', k_geom = 10)
bank <- ScaleBanksy(bank)
bank <- Banksy:::RunPCA(bank, lambda = 0, npcs = 20)
bank <- Banksy:::RunUMAP(bank, lambda = 0, npcs = 20, nneighbors = 50)
set.seed(42)
bank <- ClusterBanksy(bank, lambda = 0, pca = TRUE, npcs = 20,
                      method = 'leiden', k.neighbors = 50, resolution = 1.5)
bank_kNN_r <- copy(bank)
bank_kNN_r2 <- copy(bank)
bank_kNN_r@meta.data<-cbind(bank@meta.data, clust_kNN_r)
bank_kNN_r<-ConnectClusters(bank = bank_kNN_r, map.to = 'clust_lam0.25_k50_res1.5_kNN_r_kg10')
bank_kNN_r2@meta.data<-cbind(bank@meta.data, clust_kNN_r2)
bank_kNN_r2<-ConnectClusters(bank = bank_kNN_r2, map.to = 'clust_lam0.25_k50_res1.5_kNN_rn_kg10')
```


## Plot the spatial plots for different k_geoms, for the $1/r$ kernel

```{r, fig.height=18, fig.width=20}
num_clusters<-max(bank_kNN_r@meta.data[clust.names(bank_kNN_r)])
cols<-Banksy:::getPalette(num_clusters)
names(cols)<-1:num_clusters
cnms<-c(clust.names(bank_kNN_r)[2:length(clust.names(bank_kNN_r))], clust.names(bank_kNN_r)[1])
spatdims<-vector(mode = "list", length = length(cnms))
for (i in 1:length(cnms)){
  runid<-cnms[i]
  spatdims[[i]]<-plotSpatial(bank_kNN_r, type = 'discrete',
                             by = runid, 
                             col.discrete = cols,
                             pt.size = 0.25, 
                             main.size = 10)+facet_wrap(~feature)+ ggtitle(runid)
  
    
  
}
do.call("grid.arrange", c(spatdims, ncol = 2))


```
Alluvial plot showing the relationship between the clusterings. 
```{r, fig.height=7, fig.width=15}
plotAlluvia(bank_kNN_r)

```


## Plot the spatial plots for different k_geoms, for the $1/r^2$ kernel

```{r, fig.height=18, fig.width=20}
num_clusters<-max(bank_kNN_r2@meta.data[clust.names(bank_kNN_r2)])
cols<-Banksy:::getPalette(num_clusters)
names(cols)<-1:num_clusters
cnms<-c(clust.names(bank_kNN_r2)[2:length(clust.names(bank_kNN_r2))], clust.names(bank_kNN_r2)[1])
spatdims<-vector(mode = "list", length = length(cnms))
for (i in 1:length(cnms)){
  runid<-cnms[i]
  spatdims[[i]]<-plotSpatial(bank_kNN_r2, type = 'discrete',
                             by = runid, 
                             col.discrete = cols,
                             pt.size = 0.25, 
                             main.size = 10)+facet_wrap(~feature)+ ggtitle(runid)
  
    
  
}
do.call("grid.arrange", c(spatdims, ncol = 2))


```

Alluvial plot showing the relationship between the clusterings. 
```{r, fig.height=7, fig.width=15}
plotAlluvia(bank_kNN_r2)
# bank_test<-copy(bank_kNN_r2)
# df <- meta.data(bank_test)[, clust.names(bank_test), drop = FALSE]
# max.cells = 500; seed = 42; flow.colors = NULL
# 
# if (ncol(df) < 2) {
#   stop("Alluvia will only be plotted for at least 2 clustering runs.")
# }
# if (!is.null(max.cells)) {
#   df <- df[sample(seq_len(nrow(df)), max.cells), ]
# }
# cell <- rep(seq_len(nrow(df)), times = ncol(df))
# run <- rep(names(df), each = nrow(df))
# cluster <- factor(unlist(df))
# plotdf <- data.frame(cell = cell, run = run, cluster = cluster)
# nclust <- length(levels(cluster))
# 
# flow.colors <- Banksy:::getPalette(nclust)
# 
# StatStratum <- ggalluvial::StatStratum
# alluvia <- ggplot(plotdf, aes(axis1 = clust_lam0_k50_res1.5, 
#                               axis2 = clust_lam0.25_k50_res1.5_kNN_rn_kg5, 
#                               axis3 = clust_lam0.25_k50_res1.5_kNN_rn_kg10, 
#                               axis4 = clust_lam0.25_k50_res1.5_kNN_rn_kg20, 
#                               axis5 = clust_lam0.25_k50_res1.5_kNN_rn_kg50, 
#                               axis6 = clust_lam0.25_k50_res1.5_kNN_rn_kg100, 
#                               stratum = cluster, 
#                               alluvium = cell, fill = cluster, label = cluster)) + 
#   ggalluvial::geom_flow(stat = "alluvium", lode.guidance = "forward", 
#             color = "transparent", alpha = 0.4) + ggalluvial::geom_stratum(alpha = 0.75) + 
#   theme_minimal() + theme(panel.grid = element_blank(), 
#                           axis.line = element_line(color = "grey40"), axis.text.x = element_text(angle = 90), 
#                           axis.title.x = element_blank()) + scale_x_discrete(expand = c(0.05, 
#                                                                                         0.05)) + scale_fill_manual(values = flow.colors)
# 
# print(alluvia)
```
## 

## Session information

<details>

```{r, sess}
sessionInfo()
```

</details>
